---
title: "Light Intensity Figure"
author: "Abbi Berg"
date: "4/9/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

For the purposes of this project, would like to produce a figure that is effective at displaying how increasing light intensity increases the frequency of male-male courtship behavior in drosophila, whereas male-female courtship frequency is much less effected by the intensity of light.

# Uploading and modifying data format
## Converting Intensity_dat to long format
Code to change data formate was provided by Ethan Bahl on April 18th, 2021

```{r}

## Load data mining software and light intensity data
library(tidyverse)
Intensitydat <- read.csv("data/04112021LightProtocol.csv", )

## Reformat data for analysis
# remove blank rows and columns added by Excel.
Intensitydat <- Intensitydat[c(1:20), c(1:14)]

# remove two rows with no data
Intensitydat <- Intensitydat[-c(3:4),]

# add lux value to last few columns
colnames(Intensitydat)[12:14] = paste0(colnames(Intensitydat)[12:14], ".lux.17000")

# the lux.400 column names need to be changed to the same format
colnames(Intensitydat)[3:5] = c("wing.extension.lux.400", "chasing.lux.400", "chaining.lux.400")

## Reshape data

# create columns lux, behavior, count. 
Intensitydat_long <- pivot_longer(Intensitydat, 
                                  cols = wing.extension.lux.400:chaining.lux.17000,
                                  names_to = c("behavior", "lux"),
                                  names_pattern = "(.*).[L.   l]ux.([0-9]+)",
                                  values_to = "count")

# convert lux data to numberic data
Intensitydat_long$lux = as.numeric(Intensitydat_long$lux)
# View(Intensitydat_long)
```

# Calculating statistics for intensity data
Bars on for the bar graph portion will be ploted using means. The bar graph will
also have bars to represent the standard deviation. 

```{r}
## Function to calculate the mean and the standard deviation for each group 
# Code extracted from STHDA website on April 20th, 2021 (http://www.sthda.com/english/wiki/ggplot2-error-bars-quick-start-guide-r-software-and-data-visualization)

# data : a data frame
# varname : the name of a column containing the variable
  #to be summariezed
# groupnames : vector of column names to be used as
  # grouping variables

# producing the function for statistical analysis
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

# convert converting data variables to factor variables
Intensitydat_long$lux <- as.factor(Intensitydat_long$lux)
Intensitydat_long$MM.MF <- as.factor(Intensitydat_long$MM.MF)
Intensitydat_long$behavior <- as.factor(Intensitydat_long$behavior)

# get means and averages from data
df2 <- data_summary(Intensitydat_long, varname="count", 
                    groupnames=c("behavior", "lux","MM.MF"))

# df2 is summarized data from Intensitydat_long containing the means and std for each behavior, light intensity (lux) and for male only and male female trails (MM.MF)

# View(df2)

```

## Visualizing light intensity data
```{r}
# load ggplot
library(ggplot2)

## Make the bar graph with SEM lines as well as scatter plot of raw data points
p2 <- ggplot(NULL, aes(x= lux, y=count, fill=MM.MF)) +  #Use ggplot to set structure of the      graph. Using NULL allows you to call 2 different data frames within the same figure
  
  geom_bar(data = df2, stat="identity", color="black",  #makes the bar graph from the Intensity data means (df2 file). 
           position=position_dodge(.9)) + # position dodge allows for two categories of data (MM.MF) at the same lux value.
  
  geom_point(data = Intensitydat_long, position=position_jitterdodge(.7)) + 
  facet_grid(vars(behavior)) + # makes the raw data points from the unsummarized data set. 
  
geom_errorbar(data = df2,aes(ymin=count, ymax=count+sd), width=.2, position=position_dodge(.9)) # makes the error bars on the bar graphs using standard deviation 


## Add title and scale
p2 + labs(x="Light Intesntiy (lux)", y = "Number of Events")

print(p2)

```
